\chapter{Graphics Processing Units} \label{chapter:graphics_processing_units} 

\section{Architecture} \label{section:graphics_processing_units:architecture}
% Start by saying each GPU has an associated CPU
% Talk about treads and blocks
% thread warps, branching
% Talk about saturation, SEMs
% How it doesn't work well with adaptivity usually

\section{Data structure} \label{section:graphics_processing_units:data_structure}
% Mesh design, unstructured
% Talk about dynamic memory
% Talk about data structure? The fact we use unstructured meshes, are the worse case for GPU
% Talk about how data is allocated

\section{Process Parallelism} \label{section:graphics_processing_units:process_parallelism}
% GPU partitions
% Say something about worker processes, to be general with cpu workers.

When trying to solve very large problems, a single node with a single GPU will have insufficient
processing power and memory to solve the problem in a reasonable time, or at all. To work with these
larger problems, the work needs to be split at another level than GPU parallelism.  

The program can use multi-block meshes, with each block being worked on by one process and one GPU.
If there are fewer GPUs available on a system than there are processes, some processes will share a
single GPU by using asynchronous execution streams. The different processes communicate together
using \textit{Message Passing Interface (MPI)} at their boundaries. Since the solution data resides
on the GPU, it must first be copied to the main \textit{Central Processing Unit (CPU)} memory before
it is sent through the MPI runtime. It then needs to be copied to the receiving process' GPU. This
exchange necessitates multiple transfers on different levels and should be optimised as much as
possible. An approach to minimise the number of interfaces between the different mesh blocks is
explained in Chapter~\ref{chapter:load_balancing}.

\section{Implementation} \label{section:graphics_processing_units:implementation}
% Talk about reductions
% Talk about kernels
% Talk about reduce operations, data transfers for boundaries
