\chapter*{Acknowledgements}

I would like to thank my supervisor, Prof.\ Catherine Mavriplis. Her boundless support and patience
helped me when I felt stuck, from tiny bugs to bigger setbacks. Her insight and knowledge helped me
immensely approach new problems in ingenious ways. I am very thankful to have had the chance to have
her as my thesis supervisor.

I would also like to thank three other members of our research team. Shiqi He, François Lepage and
Mayank Vadsola overlapped with me for over a year, and acted a bit like my university of Ottawa big
brothers and sisters. They gave me no end of advice and wisdom as people who went through the same
process.

Then, I would like to thank my soon-to-be wife Ari, who always encouraged me no matter the rate at
which I was progressing. She even quickly learned new skills to help me work on this, I am very
lucky to have her by my side.

Also, I would like to thank my family. They have shown a lot of support for me, and a lot of
interest for my work. This has helped encourage me to keep progressing.

I would like to thank every teacher I had over my whole academic journey. They helped me learn a
tremendous amount, and cultivated my curiosity for the infinite number of things to learn next.

Finally, I would like to thank the different programs that supported this research. The support from
the Natural Sciences and Engineering Research Council of Canada Discovery Grant (RGPIN-2017-05320)
and a Grant from the Collaborative Research and Training Experience (CREATE- 481695-2016) program in
\acrfull{acr:SBES} are greatly appreciated. The Compute Canada organisation also helped me by
providing access to \acrshort{acr:HPC} platforms I wouldn't have been able to use otherwise, and the
necessary guidance and support to learn to use them.